---
---

## Connecting to a Database

We are going to look at the Portal mammals data organized in an SQLite database.
This is stored in a single file `portal.sqlite` that can be explored using SQL
or with tools like [SQLite Viewer](https://inloop.github.io/sqlite-viewer/) or [DB Browser](https://sqlitebrowser.org/). We will use R functions in the DBI package 
to interact with the database. 
{:.notes}


===

## Connections

The first step from RStudio is creating a connection object that
opens up a channel of communication to the database. 

```{r, handout = 0, eval = TRUE}
library(DBI)
con <- dbConnect(RSQLite::SQLite(), 
                 "data/portal.sqlite")

```


===

With the connection object availble, you can begin exploring the database.

```{r}
dbListTables(con)
```

===

```{r}
dbListFields(con, 'species')
```

===

Read an entire database table into an R data frame with `dbReadTable`, or if you
prefer "tidyverse" functions, use the [dplyr](){:.rlib} `tbl` function.

```{r, handout = 0, message=FALSE}
library(dplyr)
species <- tbl(con, 'species')
```
```{r}
species
```

===

The `dbWriteTable` function provides a mechanism for uploading data, as long as
the user specified in the connection object has permission to create tables.

```{r, handout = 0}
df <- data.frame(
  id = c(1, 2),
  name = c('Alice', 'Bob')
)
dbWriteTable(con, 'observers', df,
             append = TRUE)
```

===

```{r}
tbl(con, 'observers')
```
